Predicting the effects of character displacement on food-web dynamics

### 2019-03-18 manuscript_ECD_model.Rmd
format title page for ecology letters

$^\ast$ Corresponding author; e-mail: matthew.barbour@ieu.uzh.ch

@@

*Manuscript elements*: Figure 1, figure 2, and figure 3. All figures should be printed in color.

@@

*Keywords*: competition; eco-evolutionary dynamics; consumer-resource interactions; adaptation; ecological community; stability

@@

*Number of words*: `r wordcountaddin::word_count("manuscript_ECD_model.Rmd")`

@@

*Manuscript type*: Note. <!-- consider e-Article, depending on length -->

\footnotesize Prepared using an *Am. Nat.* inspired \LaTeX{} template for Rmarkdown. \normalsize

@@

Ecological character displacement is an important adaptive process in generating biodiversity [@Schluter2000; @Pfennig2010; but see @Stuart2013]. 
This process results from "phenotypic evolution in a species generated or maintained by [exploitative] resource competition with one or more coexisting species” [@Schluter2000]. 
Over the past 40 years, a large body of theoretical [e.g. @Lawlor1976; @Abrams1986; @Doebeli1996; @Taper1985; @McPeek2019] and empirical [reviewed in: @Schluter2000; @Dayan2005; @Stuart2013] work has examined which scenarios lead to phenotypic divergence or convergence of competing consumers.
The general result has been that, if resources are nutritionally substitutable [@Abrams1987; @Fox2008] and there is no other strong source of abundance dependence acting on consumers [@Abrams1986], then resource competition drives the adaptive divergence of competitors [@Lawlor1976; @Taper1985]. 
This adaptive process is not simply a response to statitc differences in resource distributions, but creates an eco-evolutionary feedback that drives further differentiation. 
This crucial insight was made by theoretical models that explicitly included resource dynamics as a mediator of competition in driving evolutionary change [@Lawlor1976; @Abrams1986; @Taper1985].

@@

# Methods

@@

where $r_i$ represents the intrinsic growth rate of resource $i$, $K_i$ represents the carrying capacity of resource $i$, $e_{ji}$ represents the conversion efficiency of resource $i$ into consumer $j$, and $m_j$ represents the mortality rate of consumer $j$. $F_{ji}(R_i)$ represents consumer $j$'s feeding rate (i.e. functional response) on resource $i$. 
This model is a useful characterization of a scenario where consumers compete for two distinct resources (e.g. zooplankton and benthic invertebrates) rather than if resources are better characterized by a continuous trait distribution (e.g. seed size, see @Taper1985 for an example). 
<!--
In the simplest scenario, consumers can forage for both resources simultaneously and their feeding rates increase linearly with resource abundance, such that $F_{ji}(R_i)=a_{ji}R_i$ where $a_{ji}$ is the attack rate of consumer $j$ on resource $i$ [@MacArthur1972]. 
The evolution of consumer attack rates in this model (and several extensions) have been analyzed in detail [@Lawlor1976; @Abrams1986], with the general result being divergent character displacement. 
I say consumers have undergone divergent character displacement if there evolved attack rates are more specialized, measured as $|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|$, when evolving with a competing consumer. 
-->
To investigate the ecological consequences of divergent character displacement, I analyzed differences in resource abundances and food-web stability at equilibrium. 
An equilibrium is reached when the four rates of change in the above equations are 0, and solving the system at this point gives equilibrium abundances for each resource ($\hat R_1,\hat R_2$) and consumer ($\hat C_1,\hat C_2$). 
At the coexistence equilibrium (i.e., where $\hat R_1,\hat R_2, \hat C_1,\hat C_2 > 0$), I calculated resource abundances and performed a linear stability analysis [@Otto2007]. 
This stability analysis derives the dominant eigenvalue, $\lambda$, which determines whether (and how readily) the food web will return to equilibrium after a small perturbation.

When possible, I derived analytical expressions for the relationship between consumer attack rates and food-web dynamics. 
To do this, I simplified the model by assuming that resources are equivalent ($r=r_1=r_2$ and $K=K_1=K_2$) as well as consumers ($e=e_{11}=e_{12}=e_{21}=e_{22}$; $m=m_1=m_2$), except that consumer attack rates are mirror images of each other ($a_{ii}=a_{11}=a_{22}$; $a_{ij}=a_{12}=a_{21}$). 
Controlling for other sources of variability allowed me to isolate the general effects of divergent character displacement on food-web dynamics.

Although divergent character displacement occurs in all scenarios we examined, the magnitude of this divergence may be context dependent.
Therefore, I simulated the eco-evolutionary feedback between consumers and resources using an Adaptive Dynamics approach.
At each evolutionary time step, I created a mutant consumer by randomly choosing a consumer and modifying its attack rate on one resource by either subtracting or adding a small constant (0.001 in the following simulations) with equal probability. 
The mutant's attack rate on the other resource was determined by a tradeoff, such that $\big(a_{ii}/{A}\big)^n+\big(a_{ij}/{A}\big)^n=1$, where $A$ is the total investment in attack rates and $n$ describes the shape of the tradeoff [@Sargent2006]. 
This function has the useful property that it differentiates between cases where intermediate combinations of $a_{ii}$ and $a_{ij}$ are higher, on average, than the extremes (when $n>1$, green line in Fig. \ref{fig:MacArthur_Plot}A) or, conversely, where the two extremes are higher, on average, than intermediate investments (when $n<1$, orange line Fig. \ref{fig:MacArthur_Plot}A). 
When $n=1$, the tradeoff function is linear, and all combinations of $a_{ii}$ and $a_{ij}$ have the same total attack rate (blue line in Fig. \ref{fig:MacArthur_Plot}A). 
Assuming the mutant consumer was rare, I then determined whether the mutant had higher relative fitness than the resident consumer, and thus could invade and replace the resident consumer. 
If the mutant was able to invade, I updated the attack rate of the resident consumer to the mutant attack rate and allowed consumer and resource dynamics to reach a steady state. 
I then repeated the simulation up to 10,000 times, which was sufficient for consumers to either reach an evolutionary stable strategy [ESS, @Smith1973] or an evolutionary limit (e.g. $\frac{a_{ii}}{a_{ii}+a_{ij}}$ is constrained to a maximum of 1 and minimum of 0).

@@

tradeoff_plot <- bind_rows(concaveDN_df, linear_df, concaveUP_df) %>%
  ggplot(., aes(x=aii, y=aij, color = factor(n))) + 
  geom_line() + 
  geom_point(data = initial_points, aes(x=a11, y=a12)) +
  xlab(expression(a[ii])) +
  ylab(expression(a[ij])) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)

tradeoff_plot
```

@@

<!-- I haven't made clear that this explores a specific scenario, MacArthur's model -->

In the simplest scenario, consumers can forage for both resources simultaneously and their feeding rates increase linearly with resource abundance, such that $F_{ji}(R_i)=a_{ji}R_i$ where $a_{ji}$ is the attack rate of consumer $j$ on resource $i$ [@MacArthur1972]. 
The evolution of consumer attack rates in this model (and several extensions) have been analyzed in detail [@Lawlor1976; @Abrams1986], with the general result being divergent character displacement. 
I say consumers have undergone divergent character displacement if there evolved attack rates are more specialized, measured as $|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|$, when evolving with a competing consumer.

When both consumers and resources are present, the abundance of resources at equilibrium are equivalent ($\hat R = \hat R_1 = \hat R_2$) and are determined by the following equation (derivation in supplementary material):

@@

```{r MacArthur_Resources, include=FALSE, fig.cap="\\label{fig:MacArthur_Resources}Effect of character displacement on total attack rates (A) and resource abundances (B) when consumers can forage for both resources simultaneously."}
Mac_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
  #geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(a[ii]+a[ij])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

Mac_resources <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") +
  #geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(R[i]+R[j])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

@@

```{r Stability, include=FALSE, fig.cap="\\label{fig:Stability}Effect of character displacement on local stability when consumers can (A) and cannot (B) forage for both resources simultaneously. Small and large points correspond to initial and evolved strategies, respectively."}
Stability_Mac <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) + #size=5,
  geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  ylab(expression("Stability "(-lambda[max]))) + #ylab(expression(-1%*%"Re"(lambda[max]))) +
  scale_shape_manual(values=c(24,21))  +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  geom_hline(yintercept=0, linetype="dotted", color="grey") +
  guides(fill = guide_legend(override.aes=list(shape=21)))

Stability_LS <- cbind.data.frame(select(tradeoff_df, n, aii, aij), LawlorSmith_2C_2R_tradeoff_dynamics) %>%
  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) +
  geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  ylab(expression("Stability "(-lambda[max]))) + #ylab(expression(-1%*%"Re"(lambda[max]))) +
  scale_shape_manual(values=c(24,21))  +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  geom_hline(yintercept=0, linetype="dotted", color="grey") +
  guides(fill = guide_legend(override.aes=list(shape=21)))

@@

```{r MacArthur_Plot, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:MacArthur_Plot}**Effect of character displacement when consumers can forage for both resources simultaneously.** Panel (A) shows the different tradeoffs in consumer attack rates used in the simulation (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85). Panel (B) shows how the shape of this tradeoff affects the magnitude of character displacement and the evolution of total attack rates in the consumer community. Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation. Panels (C) and (D) show how different tradeoffs indirectly affect total resource abundances (C) and food-web stability (D). These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2. We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*. We set initial consumer and resource abundances as: *R~i~* = *R~j~* = 2; *C~i~* = *C~j~* = 1."}
knitr::include_graphics("Figure_ECD_MacArthur.pdf")
```

## Consumers *cannot* forage for both resources simulanteously

The previous model, although the starting point for many models of competition, does not reflect many food webs where consumers are mobile and their foraging behavior links resources that occur in distinct habitats [@McCann2005]. 
The only character displacement model that I am aware of that examined the effect of resources in different habitats was studied by @Lawlor1976. This model is the same as before, except now the consumer's feeding rate takes the form:

@@

To examine the ecological consequences of this divergence, I again derive the equilibrium solution when both consumers and both resources are present. As before, resource abundances are equivalent at this equilibrium ($\hat R = \hat R_1 = \hat R_2$ when both consumers and resources present), but are now determined by the following equation (derivation in supplementary material):

@@

```{r LS_Resources, include=FALSE, fig.cap="\\label{fig:LS_Resources}Effect of character displacement on effective attack rates (A) and resource abundances (B) when consumers cannot forage for resources simultaneously."}
LS_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
  #geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(w*a[ii]+(1-w)*a[ij])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

>>>>

$^\ast$ Corresponding author; e-mail: matthew.barbour@ieu.uzh.ch, phone: +41 44 635 57 11

@@

*Running title*: Character divergence destabilizes food webs

@@

*Keywords*: competition; eco-evolutionary dynamics; consumer-resource interactions; adaptation; community stability

@@

*Manuscript type*: Letter. <!-- consider e-Article, depending on length -->

@@

*Number of words in abstract*:

*Number of words in main text*:

*Number of references*:

*Number of figures*:

@@

Ecological character displacement is an important adaptive process in generating biodiversity [@Schluter2000; @Pfennig2010; but see @Stuart2013]. 
This process results from "phenotypic evolution in a species generated or maintained by [exploitative] resource competition with one or more coexisting species” [@Schluter2000]. 
Over the past 40 years, a large body of theoretical [e.g. @Lawlor1976; @Abrams1986; @Doebeli1996; @Taper1985; @McPeek2019] and empirical [reviewed in: @Schluter2000; @Dayan2005; @Stuart2013] work has examined which scenarios lead to phenotypic divergence or convergence of competing consumers.
The general result has been that, if resources are nutritionally substitutable [@Abrams1987; @Fox2008] and there is no other strong source of abundance dependence acting on consumers [@Abrams1986], then resource competition drives the adaptive divergence of competitors [@Lawlor1976; @Taper1985; \ref{fig:fig_1}]. 
This adaptive process is not simply a response to statitc differences in resource distributions, but creates an eco-evolutionary feedback that drives further differentiation. 
This crucial insight was made by theoretical models that explicitly included resource dynamics as a mediator of competition in driving evolutionary change [@Lawlor1976; @Abrams1986; @Taper1985].

@@

```{r fig_1, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:fig_1}**Divergent character displacement.** A textbook example of divergent character displacement is threespine stickleback in small coastal lakes of British Columbia. Sticklback consume two main resources, zooplankton and benthic invertebrates. In isolation, stickleback evolve a generalist phenotype that can take advantage of both resources. In the presence of a competitor, however, stickleback evolve specialized phenotypes that enable them to capture different resources more efficiently. A large body of theoretical work has shown that resource competition generally drives the adaptive divergence of consumers onto different resources. Currently though, we lack theoretical predictions for how this adaptive divergence affects consumer-resource interactions and food-web dynamics."}
knitr::include_graphics("ECD_model_fig_1.pdf")
```

# Material and methods

## Underlying Consumer-Resource Dynamics

@@

where $r_i$ represents the intrinsic growth rate of resource $i$, $K_i$ represents the carrying capacity of resource $i$, $e_{ji}$ represents the conversion efficiency of resource $i$ into consumer $j$, and $m_j$ represents the mortality rate of consumer $j$. $F_{ji}(R_i)$ represents consumer $j$'s feeding rate (i.e. functional response) on resource $i$. 
This model is a useful characterization of a scenario where consumers compete for two distinct resources (e.g. zooplankton and benthic invertebrates) rather than if resources are better characterized by a continuous trait distribution (e.g. seed size, see @Taper1985 for an example).

## Foraging Scenarios

I studied three different foraging scenarios. 
In the first, I assume that consumers can forage for both resources simultaneously and their feeding rates increase linearly with resource abundance, such that:

$$F_{ji}(R_i)=a_{ji}R_i$$

where $a_{ji}$ is the attack rate of consumer $j$ on resource $i$ [@MacArthur1972]. 
This first scenario is the starting point for many models of resource competition; however, it does not reflect many food webs where consumers are mobile and their foraging behavior links resources that occur in different habitats [@McCann2005]. 
A simple extension of the above scenario takes the form:

where $w_{ji}$ represents the proportion of time consumer $j$ spends foraging in a habitat where only resource $j$ is found (i.e. habitat preference). 
Note that since $w_{ji}$ is a proportion that $w_{jj}=1-w_{ji}$. 
Finally, it is well known that consumer feeding rates often saturate at high resource abundances [@Holling1959; @Rosenzweig1963; @Murdoch2003; @McCann2011] and they do not usually spend a fixed proportion of time in a particular habitat [@McCann2005]. 
To account for these biological realities, I consider a more realistic functional response [derived by @McCann2005]:

where consumer $i$'s feeding rate on resource $i$ is influenced by the abundance of each resource; saturates as resource abundances increase (due to handling time $h$); and consumer habitat preferences are modified by the relative abundance of resources, such that: $W_{ii}=\frac{wR_i}{wR_i+(1-w)R_j}$.

Previous studies have analyzed the evolution of consumer attack rates in the first two foraging scenarios using an Adaptive Dynamics approach [@Otto2007], with the general result being divergent character displacement [@Lawlor1976; @Abrams1986]. 
I say consumers have undergone divergent character displacement if there evolved attack rates are more specialized, measured as $|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|$, when evolving with a competing consumer.
I also used an Adaptive Dynamics approach to analyze character displacement in the third foraging scenario, and I too observed divergent character displacement (detailed analysis in Appendix S1 of Supplementary Information).

```{r fig_2, echo=FALSE, fig.pos="H", fig.cap="\\label{fig:fig_2}**Foraging scenarios and evolutionary tradeoffs.** I examined the effect of ecological character displacement on food-web dynamics under different assumptions of ecological foraging scenarios and evolutionary trade-offs in consumer foraging traits. Specifically, I examined whether these effects depended on whether resources occurred in the same habitat (top left) or in different habitats (top right). I also included a more realistic functional responses. In each of these foraging scenarios, I examined the effect of different evolutionary tradeoffs in attack rates."}
knitr::include_graphics("ECD_model_fig_2.pdf")
```

## Food-web Dynamics

Given that divergent character displacement was an expected result, I focus here on its consequences for food-web dynamics across these foraging scenarios.
To do this, I analyzed differences in resource abundances and food-web stability at equilibrium. 
An equilibrium is reached when the four rates of change in the above equations are 0, and solving the system at this point gives equilibrium abundances for each resource ($\hat R_1,\hat R_2$) and consumer ($\hat C_1,\hat C_2$). 
At the coexistence equilibrium (i.e., where $\hat R_1,\hat R_2, \hat C_1,\hat C_2 > 0$), I calculated resource abundances and performed a linear stability analysis [@Otto2007]. 
This stability analysis derives the dominant eigenvalue, $\lambda$, which determines whether (and how readily) the food web will return to equilibrium after a small perturbation.

When possible, I derived analytical expressions for the relationship between consumer attack rates and food-web dynamics. 
To do this, I simplified the model by assuming that resources are equivalent ($r=r_1=r_2$ and $K=K_1=K_2$) as well as consumers ($e=e_{11}=e_{12}=e_{21}=e_{22}$; $h=h_{11}=h_{12}=h_{21}=h_{22}$; $m=m_1=m_2$), except that consumer attack rates and their habitat preferences (if present) are mirror images of each other ($a_{ii}=a_{11}=a_{22}$; $a_{ij}=a_{12}=a_{21}$; $w_{ii}=w_{11}=w_{22}$). 
Controlling for other sources of variability allowed me to isolate the general effects of divergent character displacement on food-web dynamics.

To gain insight to the eco-evolutionary feedback generated by divergent character displacement, I conducted simulations using an Adaptive Dynamics approach. 
<!--
The evolutionary dynamics of character displacement have been well characterized in some of the ecological scenarios examined here [@Lawlor1976; @Abrams1986] and are not the focus of this work. 
Still, 
Although divergent character displacement occurs in all scenarios we examined, the magnitude of this divergence may be context dependent.
Therefore, I simulated the eco-evolutionary feedback between consumers and resources using an Adaptive Dynamics approach.
-->
Specifically, at each evolutionary time step, I created a mutant consumer by randomly choosing a consumer and modifying its attack rate on one resource by either subtracting or adding a small constant (0.001 in the following simulations) with equal probability. 
The mutant's attack rate on the other resource was determined by a tradeoff, such that $\big(a_{ii}/{A}\big)^n+\big(a_{ij}/{A}\big)^n=1$, where $A$ is the total investment in attack rates and $n$ describes the shape of the tradeoff [@Sargent2006]. 
This function has the useful property that it differentiates between cases where intermediate combinations of $a_{ii}$ and $a_{ij}$ are higher, on average, than the extremes (when $n>1$, green line in Fig. \ref{fig:MacArthur_Plot}A) or, conversely, where the two extremes are higher, on average, than intermediate investments (when $n<1$, orange line Fig. \ref{fig:MacArthur_Plot}A). 
When $n=1$, the tradeoff function is linear, and all combinations of $a_{ii}$ and $a_{ij}$ have the same total attack rate (blue line in Fig. \ref{fig:MacArthur_Plot}A). 
Assuming the mutant consumer was rare, I then determined whether the mutant had higher relative fitness than the resident consumer, and thus could invade and replace the resident consumer. 
If the mutant was able to invade, I updated the attack rate of the resident consumer to the mutant attack rate and allowed consumer and resource dynamics to reach a steady state. 
I then repeated the simulation up to 10,000 times, which was sufficient for consumers to either reach an evolutionary stable strategy [ESS, @Smith1973] or an evolutionary limit (e.g. $\frac{a_{ii}}{a_{ii}+a_{ij}}$ is constrained to a maximum of 1 and minimum of 0).

@@

tradeoff_plot <- bind_rows(concaveDN_df, linear_df, concaveUP_df) %>%
  ggplot(., aes(x=aii, y=aij, color = factor(n))) + 
  geom_line(size=1.5) + 
  geom_point(data = initial_points, aes(x=a11, y=a12), size=5, show.legend=FALSE) +
  xlab(expression(bold(a[ii]))) +
  ylab(expression(bold(a[ij]))) +
  scale_color_manual(name="Evolutionary\ntradeoff (n)", values=cbPalette) +
  theme_cowplot(font_size = 30, line_size=2)# +
  #theme(axis.title = element_text(size = 30),
  #      axis.text = element_text(size = 20),
  #      legend.title = element_text(size = 25),
  #      legend.text = element_text(size = 20),
  #      legend.spacing = unit(1.0,'cm'))

save_plot(filename = "attack_tradeoffs.pdf", plot = tradeoff_plot, base_height = 6, base_width = 9)
```

@@

<!-- I haven't made clear that this explores a specific scenario, MacArthur's model

In the simplest scenario, consumers can forage for both resources simultaneously and their feeding rates increase linearly with resource abundance, such that $F_{ji}(R_i)=a_{ji}R_i$ where $a_{ji}$ is the attack rate of consumer $j$ on resource $i$ [@MacArthur1972]. 
The evolution of consumer attack rates in this model (and several extensions) have been analyzed in detail [@Lawlor1976; @Abrams1986], with the general result being divergent character displacement. 
I say consumers have undergone divergent character displacement if there evolved attack rates are more specialized, measured as $|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|$, when evolving with a competing consumer.
-->

In this first scenario, the abundance of resources at equilibrium are equivalent when both consumers and resources are present ($\hat R = \hat R_1 = \hat R_2$), and are determined by the following equation (derived in Appendix S2):

@@

```{r MacArthur_Resources, include=FALSE, fig.cap="\\label{fig:MacArthur_Resources}Effect of character displacement on total attack rates (A) and resource abundances (B) when consumers can forage for both resources simultaneously."}
Mac_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
  #geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(a[ii]+a[ij])) +
  #xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

Mac_resources <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") +
  #geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(R[i]+R[j])) +
  #xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

@@

```{r Stability, include=FALSE, fig.cap="\\label{fig:Stability}Effect of character displacement on local stability when consumers can (A) and cannot (B) forage for both resources simultaneously. Small and large points correspond to initial and evolved strategies, respectively."}
Stability_Mac <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) + #size=5,
  geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
 #xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  ylab(expression("Stability "(-lambda[max]))) + #ylab(expression(-1%*%"Re"(lambda[max]))) +
  scale_shape_manual(values=c(24,21))  +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  geom_hline(yintercept=0, linetype="dotted", color="grey") +
  guides(fill = guide_legend(override.aes=list(shape=21)))

Stability_LS <- cbind.data.frame(select(tradeoff_df, n, aii, aij), LawlorSmith_2C_2R_tradeoff_dynamics) %>%
  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) +
  geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  #xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  ylab(expression("Stability "(-lambda[max]))) + #ylab(expression(-1%*%"Re"(lambda[max]))) +
  scale_shape_manual(values=c(24,21))  +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  geom_hline(yintercept=0, linetype="dotted", color="grey") +
  guides(fill = guide_legend(override.aes=list(shape=21)))

@@

```{r MacArthur_Plot, include=FALSE, eval=FALSE, fig.pos="H", fig.cap="\\label{fig:MacArthur_Plot}**Effect of character displacement when consumers can forage for both resources simultaneously.** Panel (A) shows the different tradeoffs in consumer attack rates used in the simulation (green, *n* = 1.15; blue, *n* = 1; orange, *n* = 0.85). Panel (B) shows how the shape of this tradeoff affects the magnitude of character displacement and the evolution of total attack rates in the consumer community. Large circles (two consumers) and triangles (one consumer) correspond to the end points of the eco-evolutionary simulation. Panels (C) and (D) show how different tradeoffs indirectly affect total resource abundances (C) and food-web stability (D). These simulations used the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2. We set an initial value of *a~ii~* = 1.2 and *a~ij~* depended on the value of *n*. We set initial consumer and resource abundances as: *R~i~* = *R~j~* = 2; *C~i~* = *C~j~* = 1."}
knitr::include_graphics("Figure_ECD_MacArthur.pdf")
```

## Consumers *cannot* forage for both resources simulanteously
<!--
The previous model, although the starting point for many models of competition, does not reflect many food webs where consumers are mobile and their foraging behavior links resources that occur in distinct habitats [@McCann2005]. 
The only character displacement model that I am aware of that examined the effect of resources in different habitats was studied by @Lawlor1976. This model is the same as before, except now the consumer's feeding rate takes the form:

@@

To examine the ecological consequences of this divergence, I again derive the equilibrium solution when both consumers and both resources are present. 
As before, resource abundances are equivalent at this equilibrium ($\hat R = \hat R_1 = \hat R_2$ when both consumers and resources present), but are now determined by the following equation (derivation in supplementary material):
-->
In this second scenario, I again see that resource abundances are equivalent when both consumers and resources are present ($\hat R = \hat R_1 = \hat R_2$), but are now determined by the following equation (derived in Appendix S3):

@@

```{r LS_Resources, include=FALSE, fig.cap="\\label{fig:LS_Resources}Effect of character displacement on effective attack rates (A) and resource abundances (B) when consumers cannot forage for resources simultaneously."}
LS_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
  #geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(w*a[ii]+(1-w)*a[ij])) +
  #xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

LS_resources <- cbind.data.frame(select(tradeoff_df, n, aii, aij), LawlorSmith_2C_2R_tradeoff_dynamics) %>%
  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color=factor(n))) + 
  geom_line() +
  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
  #geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(R[i]+R[j])) +
  #xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))
